\LaTeX\ Author Guidelines for 3DIMPVT Proceedings

### 2012-06-19 3dimpvt/paper.tex
working on localization

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in applications
  such as virtual reality and environment mapping. Accurate texture
  mapping of these models is vital towards visualizing the data
  gathered by modeling systems and increases overall usability. The
  localization of cameras in 3D scenes often suffers from inaccuracies, resulting in visible discontinuties
  when images are directly projected onto a plane for
  texturing. Previous approaches at minimizing these discontinuities
  do not robustly handle a wide range of camera locations and angles, and often suffer from error accumulation when stitching together multiple
  images. We propose two variants on a new approach for reducing discontinuities
  during texture mapping, one tailored for optimal images, and one tailored for sub-optimal images.
\end{abstract}

@@

The aim of this paper is to present a solution for texture mapping the
3D models generated by indoor modeling systems, with specific
attention given to a human-operated system with higher localization
errors and greater variance in camera locations. The paper is
organized as follows. Section \ref{sec:backpackSystem} provides an
overview of the backpack modeling system from which data and examples
used throughout this paper originate. Section
\ref{sec:textureMappingOverview} describes the general problem of 3D
texture mapping and reviews existing approaches. Section
\ref{sec:textureMappingMethods} describes our new approach towards texture mapping, as well as unique challenges posed by the
backpack modeling system and our two approaches tailored for
walls and floors/ceilings respectively. Section \ref{sec:results}
compares results and presents conclusions.

\section{Backpack Modeling System}
\label{sec:backpackSystem}
Human-operated data acquisition systems provide unique advantages over
vehicular-mounted systems in terms of agility and
portability. Unfortunately, human-operated systems also suffer from a
lack of automation and stability, often resulting in much higher data
variance and localization error. As a result, common methods for
texture mapping often produce poor results, as examined in section
XX. Before discussing how to overcome these challenges, we will first
provide an overview of the backpack modeling system from which our
test data was obtained.

\subsection{Data Acquisition Hardware}
\label{sec:backpackHardware}
The backpack modeling system which we worked with contains five 2D
laser range scanners, two cameras, an orientation sensor, and an
IMU. The laser scanners have a 30-meter range and a 270$^{\circ}$
field of view. The two cameras are equipped with fisheye lenses,
resulting in a 180$^{\circ}$ field of view, and are mounted with one
facing left and one facing right. The IMU provides highly accurate
measurements of all 6 DOF at 200Hz, and is used as a ground truth
reference. The orientation sensor provides orientation parameters at a
rate of 180Hz.

With this backpack actively scanning, a human operator then takes
great care to walk a path such that every wall in the desired indoor
environment is passed and scanned lengthwise at least once.

\subsection{Environment Reconstruction}
\label{sec:environmentReconstruction}
Using data gathered by the onboard sensors and multiple localization
and loop-closure algorithms, the backpack is localized over its data
collection period, and a 3D point cloud of the surrounding environment
is constructed based on the laser scanner readings. Approximate normal
vectors for each point in the point cloud are then calculated by
gathering neighboring points within a small radius and processing them
using principal component analysis. These normal vectors allow for the
classification and grouping of adjacent points into structures such as
walls, ceilings, floors, and staircases. A RANSAC algorithm is then
employed to fit polygonal planes to these structured groupings of
points, resulting in a fully planar model. This planar model,
consisting of multiple 2D polygonal planes in 3D space, must then be
textured by all or some subset of the images captured by the
backpack's camera.

\subsection{Image Association}
\label{sec:imageAssociation}
Before discussing texture mapping, we (something goes here based on whether we take credit for stewarts work or not)

In order to speed up our texture mapping process, it is prudent to
reduce the amount of potential images being considered for texture
mapping each plane. The backpack modeling system takes 5
pictures/second from both cameras, resulting in each plane being
present in (and capable of being textured mapped by) thousands of
images, at a wide range of distances and angles. Thus, for the sake of
efficiency, it makes sense to associate each plane with a much smaller
list of images, preferably some set which contains the most desirable images
according to some metric. Thus we keep three criteria in mind while
selecting such a subset. First, we want our final collection of images
to be capable of covering the entirety of a plane, so as to ensure
there are no holes in our final texture. Second, we want all our
images to be taken from a relatively close distance to the plane, and
at as much of a direct, head-on angle as possible. This ensures that
the image projects squarely onto our desired plane, with minimal
inclusion of other planes, and we end with a texture with high
resolution. Third, we desire an amount of overlap between the images
chosen. This overlap provides us with a buffer in case we shift images
as discussed in Section X, and is also required for the blending
process presented in Section Y.

To meet these three criteria, we divide each plane into small triangles just a few centimers across. Triangulation is used due to the fairly complex
geometry of many of our planes, and is accomplished by using a
standard ear-clipping algorithm followed by further triangular
tesselation. For each triangle, we gather all images that can be
projected upon the entirety of the triangle, accounting for occlusion
with other planes as well, using standard ray-polygon intersection
tests. Of these images, we then select the best one
according to the heuristic in equation X, which favors closer distances
and more head-on projection angles. We also enforce that no two images
are chosen such that they were taken within 0.5 meters of eachother,
so as to reduce our overall amount of images. The end result is a
greatly-reduced set of images such that each one is objectively a good choice for texturing some number of areas on the plane. Furthermore, since cameras are only at good
angles to triangles near their center of projection, and poor
candidates for triangles near their edges, images will be chosen such
that there is high overlap between successive images. With the three criteria satisfied, we have reduced the amount of candidate images for
each plane from many thousands to around 50-100.

\subsection{Occlusion Checking}
Before we proceed with texturing our planes, we need to ensure that
each plane's associated images contain only content that should be
mapped onto the target plane in question. From the above process,
images were chosen such that they had an unoccluded view of one or
more trianglular sections upon a plane. This may not necessarily hold
true however for all areas of the plane, as demonstrated in figure
SOMETHING. If images are not modified to mask out occluded areas,
different images could attempt to cover the same location on a plane
with drastically different textures, inevitably leading to
undesireable discontinuities in our final texture.

Fortunately, by virtue of our indoor environments, the vast majority
of surface geometry is either horizontal or vertical, with high
amounts of right angles. This means that the areas we wish to mask out
of our images due to occlusion are nearly always rectangular
areas. Similar to the triangular tesselation performed for image
association above, we split each image into rectangular pieces, and
perform occlusion checks for a sampling of points within each
piece. For rectangles with occluded points, we then subdivide them
into smaller rectangles, and repeat the process, until we reach some
manually-set recursion limit. To actually occlude out rectangles, we
simply remove their texture, as untextured areas will never be chosen
for texture mapping. We are now ready to begin the texture mapping
process.

\section{Texture Mapping Overview}
\label{sec:textureMappingOverview}

@@

In this section, we will discuss the process of texture mapping a
single plane, as the texturing of each of our planes is completely
independent and can be completed in parallel.

The geometry of the general texture mapping process is shown in Figure 1.

 As
described in the previous section, we are provided with a set of $M$
images for each plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point in
image $i$'s coordinates. A camera matrix $P_i$ is composed of the
camera's intrinsic parameters, such as focal length and image center,
as well as the extrinsic parameters which specify the rotation and
translation of the camera's position with respect to the world
coordinates at the time that image $i$ was taken. These extrinsic
parameters are determined by the localization hardware and algorithms
mentioned in Section \ref{sec:backpackSystem}. A point $X$ on the
plane in 3D space can be related to its corresponding pixel $x$ in image $i$
through the following equation:

@@

For the sake of simplicity, we treat all planes as rectangles by generating a minimum bounding box for them. Their original vertices can be recovered later as needed, and used to crop any excess texture. A plane to be textured is thus defined by its bounding box with corners $C_i$ in
world coordinates and a normal vector indicating the front facing side
of the plane. Our goal is to texture this plane using its associated
images, while eliminating any visual discontinuities or seams that
would suggest that the plane's texture was not composed of a single
continuous image.

\subsection{Naive Mapping}
\label{sec:naive}
\begin{figure}
  \centering
  \includegraphics[height=1.5in]{naive.png}
  \caption{The result of naive texture mapping based on the imprecise
    camera matrices estimated by the localization system.}
  \label{fig:naive}
\end{figure}

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
one can texture map the plane by simply applying textures according to
the image association process and scoring function described in
Section X. This naive mapping approach independently selects the best
texture for each trianglular section on the plane, but it also means
that all of our $M$ images are used for the overall texture, thus
increasing the potential amount of poor image boundaries.

As Figure \ref{fig:naive} demonstrates, this approach leads to many
image boundaries with abrupt discontinuities, due to significant
misalignment between images. One approach also tuned for our datasets
attempted to minimize these discontinuities through more intelligent
image selection, making use of spatial and temporal caching
mechanisms.

\subsection{Mapping with Caching}
In order to reduce the severity of discontinuities between adjacent
triangles, one can texture them with the same image, if possible,
eliminating the discontinuity outright, or by two images taken from
similar positions, such that any texture mismatch is ideally less
pronounced. This is accomplished by adding a 2-level cache to the
native mapping approach, and sorting triangles in an order such that
triangle $t_i$ is adjacent to triangle $t_{i+1}$. When a
triangle is assigned a texture, the image it selected is added to the
level 1 cache, and other images that were taken nearby in time and
location, within a predefined threshold, are added to the level 2
cache. Each subsequent triangle to be textured then searches through
the level 1 cache, selecting the best image according to the same
scoring function, and comparing it to some pre-defined threshold. If
the best image's score is below this threshold, the level 2 cache is
then searched, and then finally if necessary, the set of all images is
searched. In this way, the entire plane is textured, with image reuse prioritized, followed by a preference for nearby images.

This approach is the method previously used by the modeling team
associated with our dataset. Therefore, we will use it as our baseline
when comparing other approaches.  The addition of the caching
structure reduces discontinuities overall, as can be seen in the above
image. Nevertheless, the amount of remaining discontinuities suggests
that image selection alone cannot produce seamless textures.  In order
to produce photorealistic texture mapping, either the camera matrices
need to be refined such that the localization is pixel accurate,
resulting in a perfect mapping, or image stitching techniques need to
be applied to provide this illusion. A commonly-used technique for
each approach is examined next.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying realistic textures to these models is an important
  step in generating accurate visualizations of data gathered by
  modeling systems.  Unfortunately, the localization of cameras in
  such systems often suffer from inaccuracies, resulting in visible
  discontinuties when different images are projected adjacently onto a
  plane for texturing. Previous approaches at minimizing these
  discontinuities do not robustly handle a wide range of camera
  locations and angles, and often suffer from error accumulation when
  stitching together multiple images. We propose two approaches for
  reducing discontinuities during texture mapping, one to robustly
  accomodate images of all orientations, and one to take advantage of
  optimal situations where images have uniform orientation.
\end{abstract}

@@

The aim of this paper is to present a solution for texture mapping the
3D models generated by indoor modeling systems, with specific
attention given to a human-operated system with higher localization
errors and greater variance in camera locations. The paper is
organized as follows. Section 2 provides an overview of the backpack
modeling system and data processing pipeline from which input data and
examples used throughout this paper originate. Section 3 describes the
general problem of texture mapping and examines two simple mapping
approaches and their problems.  Section 4 demonstrates previous
attempts at localization refinement, and presents our new approach to
image alignment. Section 5 combines our image alignment algorithm with
a simple image selection approach, as well as with a new method for image selection that
takes advantage of optimal projections. Section 6 contains results and
conclusions.

\section{Backpack Modeling System}
Human-operated data acquisition systems provide unique advantages over
vehicular-mounted systems in terms of agility and
portability. Unfortunately, human-operated systems also suffer from a
lack of automation and stability, resulting in much higher data
variance and localization error. As a result, common methods for
texture mapping generally produce poor results, as shown in section
3. Before discussing how to overcome these challenges, we will first
provide an overview of the backpack modeling system from which our
test data was obtained.

\subsection{Data Acquisition Hardware}
The backpack modeling system that captured our data contains five 2D
laser range scanners, two cameras, an orientation sensor, and an
IMU. The laser scanners are mounted orthogonally and have a 30-meter
range and a 270$^{\circ}$ field of view. The two cameras are equipped
with fisheye lenses, resulting in a 180$^{\circ}$ field of view, and
are mounted with one facing left and one facing right. These cameras take images at the rate of 5 Hz. The orientation
sensor provides orientation parameters at a rate of 180 Hz. The IMU
provides highly accurate measurements of all 6 DOF at 200 Hz, and is
used as a ground truth reference.

With the backpack actively scanning, the human operator wearing it
then takes great care to walk a path such that every wall in the
desired indoor environment is passed and scanned lengthwise at least
once.

\subsection{Environment Reconstruction}
Using data gathered by the onboard sensors and multiple localization
and loop-closure algorithms, the backpack is first localized over its data
collection period, and a 3D point cloud of the surrounding environment
is constructed based on the laser scanner readings relative to the backpack. Approximate normal
vectors for each point in the point cloud are then calculated by
gathering neighboring points within a small radius and processing them
through principal component analysis. These normal vectors allow for
the classification and grouping of adjacent points into structures
such as walls, ceilings, floors, and staircases. A RANSAC algorithm is
then employed to fit polygonal planes to these structured groupings of
points, resulting in a fully planar model. This model,
consisting of multiple 2D polygonal planes in 3D space, along with the
set of images captured by the backpack's camera, can be considered the
input to our texture mapping problem.

\section{Simple Texture Mapping}

@@

In all subsequent sections, we will discuss the process of texture
mapping a single plane, as the texturing of each of our planes is
completely independent and can be completed in parallel.

The geometry of the texture mapping process for a plane is shown in
Figure 1.  As described in the previous section, we are provided with
a set of $M$ images with which we must texture our target plane. Each
image has a camera matrix $P_i$ for $i=1..M$, which translates a 3D
point in the world coordinate system to a 2D point or pixel in image
$i$'s coordinates. If the 3D world point is not contained in the image, the 2D
point will simply be outside of the image boundaries. A camera matrix
$P_i$ is composed of the camera's intrinsic parameters, such as focal
length and image center, as well as extrinsic parameters which specify
the rotation and translation of the camera's position in 3D world coordinates at the time that image $i$ was taken. These
extrinsic parameters are determined by the backpack hardware and
localization algorithms mentioned in Section 2. A point $X$ on the
plane in 3D space can be related to its corresponding pixel $x$ in
image $i$ through the following equation:

@@

For the sake of simplicity, we treat all planes as rectangles by
generating minimum bounding boxes for them. Since our final textures will be
stored as standard rectangular images anyway, we can simply leave the
area between plane boundary and bounding box untextured, or crop it
out as needed. A plane to be textured is thus defined by a bounding
box with corners $C_i$ in world coordinates and a normal vector
indicating the front facing side of the plane. Our goal is to texture
this plane using images captured by the backpack, while eliminating
any visual discontinuities or seams that would suggest that the
plane's texture was not composed of a single continuous image.

\subsection{Direct Mapping}

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square
tiles, generally about 5 pixels across, and picking an image to texture that tile with. In order to
do this, we must first gather a list of candidate images for each tile
$t$. These candidate images must contain all four corners of $t$,
which we can quickly check using the projection method
above. Furthermore, each candidate image must have been taken at a
time when its camera had a clear line-of-sight to the center of $t$,
which can be calculated using standard ray-polygon intersection tests
between the camera location, our section, and other planes, all in
world coordinates.

Once we have a list of candidate images for $t$, we must define a
scoring function in order to compare images and objectively select the
best one. Since camera localization errors compound over distance, we
wish to minimize the distance between cameras used for texturing and
our plane. Additionally, we desire images that project squarely onto
the plane, maximizing the resolution and amount of useful texture
available in their projections.

A scoring function to maximize these two criteria is as follows:

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.png}
  \caption{For each tile, we want to select an image such that the
    function $abcdefg/awerw$ is maximized}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{directMapping.png}
  \caption{The result of direct texture mapping based on locally optimized textures.}
  \label{fig:directMapping}
\end{figure}

As Figure \ref{fig:directMapping} demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images.

\subsection{Mapping with Caching}
Since discontinuities occur where adjacent tiles select different
images that do not match up well, it makes sense to take into account
image choices made by neighboring tiles while selecting the best image
for a tile. By using the same image across tile boundaries, we
eliminate the discontinuity altogether. If this is not possible, using
very similar images will result in less noticeable discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate
before searching through the entire set. The first subset of images is
the set of images selected by adjacent tiles that have already been
processed. We must first check which images can map to $t$, and then
of those, we make a choice according to the same scoring function in
figure \ref{fig:scoringFunction}. Rather than blindly reusing this
image, we ensure it meets a predefined score threshold to be
considered a good image. If no good image is found, we then check our
second subset of images, which consists of images that were taken near
the images in the first subset, both spatially and temporally. These
images are not the same as the ones used for neighboring tiles, but
they were taken at a similar location and time, suggesting that their
localization and projection will be very similar. Again, if no good
image is found according to a threshold, we then must search the
entire set of candidate images.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{caching.png}
  \caption{The result of adding a caching system to locally optimized textures.}
  \label{fig:caching}
\end{figure}

The results of this caching approach are shown above in figure
\ref{fig:caching}.  Discontinuities have been reduced overall, but the
amount of remaining seams suggests that image selection
alone cannot produce seamless textures. The image projections themselves will have to be adjusted in order to reliably produce clean textures.

\section{Localization Refinement}
In order to produce photorealistic texture mapping, either camera
matrices need to be refined such that their localization is pixel
accurate, resulting in a perfect mapping, or image stitching
techniques need to be applied to provide this illusion. Before
examining these approaches, we first obtain a set of images to work
with.

Rather than perform camera or image adjustments across many thousands
of images, we choose to work with the more limited set of images
corresponding to those chosen by the direct mapping approach, without
caching. This set of images constitutes a good candidate set for
generating a final seamless texture, since it meets three important
criteria. First, this set of images contains at least one image that
covers each tile on our plane. Thus, unless extreme changes in
localization occur, we can ensure that there will be no holes in our
final texture. Second, since images are all selected according to the
same scoring function in figure \ref{fig:scoringFunction}, we know
that our images are all taken at a head-on angle and should project
onto the plane in similar ways. Third, as a side result of the scoring
function, selected images are only good candidates for the tiles near
their center of projection. Thus, there should be plenty of overlap
between selected images, allowing for some degree of shifting as well
as blending between them. With this set of images, we now examine two
approaches towards refining and combining their projections.
